### 2024-05-20 Sections/Results_Sensitivity.tex

\documentclass[../Article_Sensitivity_Analsysis.tex]{subfiles}
\graphicspath{{\subfix{../Figures/}}}
\begin{document}

	\label{CH: Results}

	This work investigates the influence of pressure change on the state space and the extraction yield. The process model and parameters have been discussed in {\color{red}article 1}. The process model was calibrated on the set of experiments obtained at different operating conditions, $30 - 40^\circ C$, $100 - 200$ bar, and $3.33-6.67 \cdot 10^{-5}$ kg/s. The sensitivity analysis has been performed assuming that the system operates at $35^\circ C$, 150 bar and 5 $\cdot 10^{-5}$ kg/s.

	%\subfile{Sensitivity_Flow_Rate}

    \subfile{Sensitivity_Pressure}

    %\texttt{}\subfile{Sensitivity_Inlet_Temperature}

\end{document}