### 2024-06-07 Sections/Results_Sensitivity.tex

\documentclass[../Article_Sensitivity_Analsysis.tex]{subfiles}
\graphicspath{{\subfix{../Figures/}}}
\begin{document}

	\label{CH: Results}

	%This work investigates the influence of pressure change on the state space and the extraction yield. The process model and parameters have been discussed in {\color{red}article 1}. The process model was calibrated on the set of experiments obtained at different operating conditions, $30 - 40^\circ C$, $100 - 200$ bar, and $3.33-6.67 \cdot 10^{-5}$ kg/s. The sensitivity analysis has been performed assuming that the system operates in the middle of validate range, hence at $35^\circ C$, 150 bar and 5 $\cdot 10^{-5}$ kg/s.

	This study examines the impact of pressure change on the state space and extraction yield. The process model and parameters are described in {\color{red}article 1}. Calibration of the process model is based on experiments performed under varying operating conditions: temperatures between $30 - 40^\circ C$, pressures between $100 - 200$ bar, and mass flow between $3.33-6.67 \cdot 10^{-5}$ kg/s. The local sensitivity analysis was conducted, assuming the system operates at the midpoint of the validated range: $35^\circ C$, 150 bar and 5 $\cdot 10^{-5}$ kg/s.

	%\subfile{Sensitivity_Flow_Rate}

    \subfile{Sensitivity_Pressure}

    %\texttt{}\subfile{Sensitivity_Inlet_Temperature}

\end{document}