### 2024-07-06 Sections/Results_Sensitivity.tex

\documentclass[../Article_Sensitivity_Analsysis.tex]{subfiles}
\graphicspath{{\subfix{../Figures/}}}
\begin{document}

	\label{CH: Results}

	The details of the process model and parameters are delivered in {\color{red}article 1}. The process model was validated under the following range of operating conditions: temperatures between $30 - 40^\circ C$, pressures between $100 - 200$ bar, and mass flow rates between $3.33-6.67 \cdot 10^{-5}$ kg/s. The sensitivity analysis is performed at the midpoint of the validated range: $35^\circ C$, 150 bar and 5 $\cdot 10^{-5}$ kg/s. This study examines the impact of pressure change on the state space and extraction yield through the local sensitivity analysis. The local sensitivity analysis results can be interpreted as the system's response to an infinitesimal deviation in pressure.

	%\subfile{Sensitivity_Flow_Rate}

    \subfile{Sensitivity_Pressure}

    %\texttt{}\subfile{Sensitivity_Inlet_Temperature}

\end{document}